---
Reproducible Research Peer Assessment 1
=======================================


First, we unzip the data file in the Github repository:

```{r, echo = "FALSE"}
unzip("rawdata.zip")
```

Read the file and format date:
```{r, echo = "FALSE"}
ActiveData <- read.csv("activity.csv")
library(lubridate)
ActiveData$date <- ymd(ActiveData$date)
```
Calculate the total number of steps taken per day
```{r}
library(dplyr)
ActiveDays <- group_by(ActiveData, date)
Sums <- summarise(ActiveDays, steps=sum(steps))
```
Create a histogram of results:
```{r}
hist(Sums$steps, main = "Histogram - Daily Sum of Steps", xlab = "Daily Sum of Steps")
```
  
Calculate and report the mean and median of the total number of steps taken per day
```{r, echo = "FALSE"}
MeanSteps <- mean(Sums$steps, na.rm = TRUE)
MeanSteps <- round(MeanSteps, digits = 0)
MedianSteps <- median(Sums$steps, na.rm = TRUE)
MedianSteps <- round(MedianSteps, digits = 0)
options(scipen=999)
```
The mean number of steps per day is `r MeanSteps`. The median is `r MedianSteps`.

Make a time series plot of the average daily activity pattern:  

```{r, fig.height = 4}
library(dplyr)
Intervals <- group_by(ActiveData, interval)
AvgInterval <- summarize(Intervals, steps=mean(steps, na.rm=TRUE))
with(AvgInterval, plot(interval, steps, type = "l", main = "Average Interval Activity Pattern",xlab = "avg. 5-min intervals over 24 hours"))
```

```{r, echo = "FALSE", results = "hide"}
max <- which.max(AvgInterval$steps)
MaxInt <- AvgInterval[max, 1]
```
The 5-minute interval, on average across all the days in the dataset,  that contains the maximum number of steps is `r MaxInt`.

```{r, echo = "FALSE", results = "hide"}
SumNA <- sum(is.na(ActiveData$steps))
```

Total number of rows with NAs is `r SumNA`.

Fill in all of the missing values in the dataset:  

```{r}
## Using Average data by 5-min interval previously calculated, replace NAs with the Average for that interval.

names(AvgInterval) <- c("interval", "impute")

##make Average Data the same length as the Activity Data

ImputeAvg <- do.call("rbind", replicate(61, AvgInterval, simplify = FALSE))

## Convert to Data Frames and replace NAs with imputed data
df.ActiveData <- data.frame(ActiveData)
df.ImputeAvg <- data.frame(ImputeAvg)
df.ActiveData$steps[is.na(df.ActiveData$steps)] <- df.ImputeAvg$impute[is.na(df.ActiveData$steps)]

```
  
Calculate the total number of steps taken per day

```{r}
ActiveDays <- group_by(df.ActiveData, date)
Sums <- summarise(ActiveDays, steps=sum(steps))
```

Create a histogram of results:  

```{r}
hist(Sums$steps, main = "Histogram - Daily Sum of Steps", xlab = "Daily Sum of Steps")
```
  
Calculate and report the mean and median of the total number of steps taken per day  

```{r, echo = "FALSE"}
MeanSteps <- mean(Sums$steps, na.rm = TRUE)
MedianSteps <- median(Sums$steps, na.rm = TRUE)
MeanSteps <- round(MeanSteps, digits = 0)
MedianSteps <- round(MedianSteps, digits = 0)
options(scipen=999)
```
  
The mean number of steps per day is `r MeanSteps`. The median is `r MedianSteps`.  

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r, echo = "FALSE"}
df.ActiveData <- within(df.ActiveData, weekday <- weekdays(date))
df.ActiveData$weekend = df.ActiveData$weekday %in% c("Saturday", "Sunday")
df.ActiveData$weekend <- ifelse(df.ActiveData$weekend == "TRUE", "weekend", "weekday")

```
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).  

```{r, echo = "FALSE"}
library(ggplot2)
WeekInterval <- group_by(df.ActiveData, interval, weekend)
AvgInterval <- summarize(WeekInterval, steps=mean(steps, na.rm=TRUE))
g <- ggplot(AvgInterval, aes(interval, steps))
g + geom_line() + facet_grid(weekend ~ .)

```

